\documentclass{article}
\usepackage{header}

\usepackage[]{geometry}


%\setlength{\voffset}{-0.5in}
\setlength{\headsep}{0pt}
\setlength{\textheight}{250mm}
%\setlength{\footskip}{0mm}
%\setlength{\textwidth}{480pt}
%\setlength{\hoffset}{-0.25in}

\begin{document}

\section*{Further Reading / Listening}

\begin{itemize}
    \item \textit{Measles; women in maths; Paula Milne on political marriages; Kerry Young} (2013), BBC Radio 4, 16/04, Available at : \url{https://www.bbc.co.uk/programmes/b01rvpk8} \\
    Between 7:16 and 16:15 the panel talk about the scarcity of female professors in UK universities. 
    
    
    \item \text{Emmy Noether} (2019), BBC Radio 4, 24/01, Available at : \url{https://www.bbc.co.uk/programmes/m00025bw} \\
    The program discusses the life and work of Emmy Noether, and provides a reading list of its own. 
    
    
    \item \text{Ada Lovelace} (2008), BBC Radio 4, 06/03, Available at : \url{https://www.bbc.co.uk/programmes/b0092j0x} \\
    Radio 4's 'In Our Time' on the life of Ada Lovelace. 
    
    
    \item J J O'Connor, E F Robertson, \textit{Pandrosion - man or woman?} (2018), MacTutor History of Mathematics, Gre \url{http://www-history.mcs.st-andrews.ac.uk/Extras/Pandrosion_man_woman.html} \\
    A discussion of the arguments for Pandrosion being a woman, and a comment on the difficulty of women being acknowledged for their work in classical works. 
    
    
    \item R Flood, \textit{Gauss and Germain} (2016), Gresham College, \\
    \url{http://www.gresham.ac.uk/lectures-and-events/gauss-and-germain} \\
    A lecture on the life of Germain and her important correspondence with Gauss. Germain is a really interesting mathematician and this resource gives a great introduction to her work. 
    
    
    \item L Riddle, \textit{Biographies of Women Mathematicians} (2019), Agnes Scott College, \url{https://www.agnesscott.edu/lriddle/women/women.htm} \\
    A database with biographies of nearly 200 women mathematicians, each with their own references and additional resources.
    
    
    \item L Kelley, \textit{Why Were So Few Mathematicians Female?} (1996), The Mathematics Teacher, 89(7), 592-596. Retrieved from http://www.jstor.org/stable/27969922 \\
    A paper giving a richer historical background into many of the mathematicians mentioned today. 
    
    
    \item K Henderson, \textit{Women in Mathematics} (2018), London Mathematical Society, \url{https://www.lms.ac.uk/womeninmaths} \\
    A large list of resources, including talks to inspire new mathematicians, improve departmental mindset, and statistics from 2015 in the gender-split in mathematics higher education.
    
    
    \item S Stacey, \textit{Women in maths: why the numbers don’t add up} (2017), Varsity, \url{https://www.varsity.co.uk/news/13945} \\
    A Varsity article written on the poor gender balance of maths undergraduates at the University of Cambridge, including comments from members of the mathematics faculty. 
    
    
    \item \textit{What do you know about the gender gap in maths?} (2018), The Open University,\\
    \url{https://www.open.edu/openlearn/science-maths-technology/mathematics-statistics/what-do-you-know-about-the-gender-gap-maths} \\
    An online quiz giving a 10 question test about the historic gender gap in mathematics. 
    
    
    \item H Mihaljević-Brandt, L Santamaría, M Tullney, \textit{The Effect of Gender in the Publication Patterns in Mathematics} (2016), PLOS ONE, 11(10), e0165367 \\
    A paper on the discrepancy in publishing statistics between men and women. 

    
\end{itemize}

\end{document}